Make the full map of Tanzania

### 2022-02-21 docs/tanzania_full_map.Rmd
Finish all work before training

- Generate guessing labels (done)
- Random select and manually check validate dataset
- Fine tune a U-Net model for the whole country

### Validate dataset generation
#### Check the quality of labels
```{r}
tiles <- read_sf("results/tanzania/catalog_sample_tiles.geojson")
library(tidyr)
# Check labels
stats_guess_labels <- do.call(
    rbind, 
    lapply(unique(tiles$tile), 
           function(tile_id){
               sub_tiles <- tiles %>% filter(tile == tile_id) %>% 
                   st_drop_geometry() %>% 
                   mutate(surfix = paste(tile, index, sep = '_')) %>% 
                   mutate(score_path = file.path(here('results/tanzania/guess_labels'),
                                                 sprintf('score_%s.tif', surfix)))
               do.call(rbind, lapply(sub_tiles$surfix, function(sur) {
                   pth <- sub_tiles %>% filter(surfix == sur) %>% 
                       pull(score_path)
                   scores <- values(rast(pth))
                   na_ratio <- sum(is.na(scores))/ length(scores)
                   data.frame(surfix = sur,
                              hasna = any(is.na(scores)), 
                              na_ratio = na_ratio,
                              mean = mean(scores, na.rm = T),
                              sd = sd(scores, na.rm = T),
                              min = min(scores, na.rm = T),
                              max = max(scores, na.rm = T))}))
               }
           )
    )
save(stats_guess_labels, file = 'results/tanzania/stats_guess_labels.rda')
err_labels <- stats_guess_labels %>% filter(hasna == TRUE) %>% 
    separate(col = surfix, into = c('tile', 'index'), sep = "_")
```

@@

There are a few problematic images

@@

#### Get candidate validation

@@

Why random ones rather than good ones? To optimize the cooperation of random forest and U-Net. Because if selecting good ones, the results might be biased to easy-to-model instances.
```{r}
skip_tile <- err_labels %>% pull(tile) %>% unique()
stats_guess_labels <- stats_guess_labels %>% 
    separate(col = surfix, into = c('tile', 'index'), sep = "_") %>% 
    filter(!tile %in% skip_tile)
set.seed(124)
validate_labels <- stats_guess_labels %>% 
    group_by(tile) %>% 
    sample_n(1) %>% ungroup() %>% 
    dplyr::select(tile, index) %>% 
    mutate(index = as.integer(index))

@@

Next, the author manually checked these validation sub-tiles. Only the good ones would be used for validation.

First, move these validation candidate into a separate folder.

```{r}
copy_from <- here("results/tanzania/guess_labels")
copy_to <- here('results/tanzania/temp')
if (!dir.exists(copy_to)) dir.create(copy_to)
tiles_validate <- tiles_validate %>% 
    mutate(tile_id = paste(tile, index, sep = "_"))
cp <- lapply(tiles_validate$tile_id, function(tile_move) {
    fnames_from <- sprintf("%s/%s_%s.tif", 
                           copy_from, c("guess", "scores", "score"), tile_move)
    fnames_to <- sprintf("%s/%s_%s.tif", 
                         copy_to, c("guess", "scores", "score"), tile_move)
    file.copy(from = fnames_from, 
              to = fnames_to)
})
```

Next step is to check and fix other labels

```{r}
tile_index_val <- tiles_validate %>% 
    mutate(tile_index = paste(tile, index, sep = "_")) %>% 
    pull(tile_index)
tile_index_train <- stats_guess_labels %>% 
    mutate(index = as.integer(index)) %>% 
    mutate(tile_index = paste(tile, index, sep = "_")) %>% 
    filter(!tile_index %in% tile_index_val) %>% 
    dplyr::select(tile, index)
tiles_train <- right_join(tiles, tile_index_train, by = c("tile", "index")) %>% 
    mutate(use = 1, modify = 0, comment = NA)
write_sf(tiles_train, "results/tanzania/catalog_tiles_train.geojson")
```

>>>>

- Generate guessing labels (done)
- Manually check labels (done)
- Train U-Net model for the whole country

## Steps
### Generate guessing labels

Prepare the needed datasets and install necessary software (e.g. GRASS GIS) and libraries. Run the following scripts under guess_model to get the guessed labels:

- 0_set_up.R
- 1_ensemble_labels.R
- 2_get_osm.R
- 3_update_lc_labels_osm.R
- 4_sampling.R
- 5_get_training.R
- 6_train_guess_model.R
- 7_pred_guess_model.R

### Manually check labels

Analyze the reliability of labels based on the score of guessed label and the quality of satellite images.

#### Check the quality of labels

```{r}
tiles <- read_sf("results/tanzania/catalog_sample_tiles.geojson")
library(tidyr)
# Check labels
stats_guess_labels <- do.call(
    rbind, 
    lapply(unique(tiles$tile), 
           function(tile_id){
               sub_tiles <- tiles %>% filter(tile == tile_id) %>% 
                   st_drop_geometry() %>% 
                   mutate(surfix = paste(tile, index, sep = '_')) %>% 
                   mutate(score_path = file.path(here('results/tanzania/guess_labels'),
                                                 sprintf('score_%s.tif', surfix)))
               do.call(rbind, lapply(sub_tiles$surfix, function(sur) {
                   pth <- sub_tiles %>% filter(surfix == sur) %>% 
                       pull(score_path)
                   scores <- values(rast(pth))
                   na_ratio <- sum(is.na(scores))/ length(scores)
                   data.frame(surfix = sur,
                              hasna = any(is.na(scores)), 
                              na_ratio = na_ratio,
                              mean = mean(scores, na.rm = T),
                              sd = sd(scores, na.rm = T),
                              min = min(scores, na.rm = T),
                              max = max(scores, na.rm = T))}))
               }
           )
    )
save(stats_guess_labels, file = 'results/tanzania/stats_guess_labels.rda')
err_labels <- stats_guess_labels %>% filter(hasna == TRUE) %>% 
    separate(col = surfix, into = c('tile', 'index'), sep = "_")
```

@@

[*NOTE:*] There are a few problematic images. Skip them for now. These tiles can be fixed after segmentation.

@@

#### Get candidate validation and train

@@

Why random ones rather than good ones? To optimize the cooperation of random forest and U-Net. Because if selecting good ones, the results might be biased to easy-to-model instances.

```{r}
# Validation
skip_tile <- err_labels %>% pull(tile) %>% unique()
stats_guess_labels <- stats_guess_labels %>% 
    separate(col = surfix, into = c('tile', 'index'), sep = "_") %>% 
    filter(!tile %in% skip_tile)
set.seed(124)
validate_labels <- stats_guess_labels %>% 
    group_by(tile) %>% 
    sample_n(1) %>% ungroup() %>% 
    dplyr::select(tile, index) %>% 
    mutate(index = as.integer(index))

@@

```{r}
# Train
tile_index_val <- tiles_validate %>% 
    mutate(tile_index = paste(tile, index, sep = "_")) %>% 
    pull(tile_index)
tile_index_train <- stats_guess_labels %>% 
    mutate(index = as.integer(index)) %>% 
    mutate(tile_index = paste(tile, index, sep = "_")) %>% 
    filter(!tile_index %in% tile_index_val) %>% 
    dplyr::select(tile, index)
tiles_train <- right_join(tiles, tile_index_train, by = c("tile", "index")) %>% 
    mutate(use = 1, modify = 0, comment = NA)
write_sf(tiles_train, "results/tanzania/catalog_tiles_train.geojson")
```

#### Manually check

Next, the author manually checked sub-tiles of validation and train. Script `8_human_refine.R` can help to do this step.

#### Add existing labels of north to use

Add to catalogs

```{r}
# Get labels from north
## Randomly get 50% tiles
## Get two for training

# Read train labels
num_tiles <- read_sf(here("data/geoms/tiles_nicfi_north.geojson")) %>% nrow()
north_train <- read.csv(here("results/north/dl_catalog_train.csv"),
  stringsAsFactors = F)

# Keep two for each tile
north_train <- do.call(
  rbind,
  lapply(
    unique(north_train$tile),
    function(this_tile) {
      north_train %>%
        filter(tile == this_tile) %>%
        arrange(desc(score)) %>%
        slice(1:2)}))

# Read validate labels
north_valid <- read.csv(here("results/north/dl_catalog_valid.csv"),
  stringsAsFactors = F)

# Randomly select 50% of tile to use
set.seed(10)
tile_to_use <- north_train %>%
  group_by(tile) %>%
  summarise(score = sum(score)) %>%
  filter(score == 10) %>%
  sample_n(round(num_tiles * 0.5)) %>%
  pull(tile)

# Finalize train and validate
train_path <- 'train'
valid_path <- 'validation'
north_train <- north_train %>%
  filter(tile %in% tile_to_use) %>%
    mutate(
    label = file.path(
      train_path,
      paste0(tile, "_", index, "_label.tif")
    ),
    img = file.path(
      train_path,
      paste0(tile, "_", index, "_img.tif")
    )) %>% 
  dplyr::select(tile_id, tile, index, label, img)
north_valid <- north_valid %>%
  filter(tile %in% tile_to_use) %>%
    mutate(
    label = file.path(
      valid_path,
      paste0(tile, "_", index, "_label.tif")
    ),
    img = file.path(
      valid_path,
      paste0(tile, "_", index, "_img.tif")
    )) %>% 
  dplyr::select(tile_id, tile, index, label, img)

# Add to catalog of full country
catalog_train <- read.csv(here("results/tanzania/dl_catalog_train.csv"),
                          stringsAsFactors = FALSE) %>% 
    rbind(north_train)
catalog_validate <- read.csv(here("results/tanzania/dl_catalog_valid.csv"),
                             stringsAsFactors = FALSE) %>% 
    rbind(north_valid)
write.csv(catalog_train, 
          here('results/tanzania/dl_catalog_train.csv'),
          row.names = F)
write.csv(catalog_validate, 
          here('results/tanzania/dl_catalog_valid.csv'),
          row.names = F)
```

Double check

```{r}
catalog_train <- read.csv(here("results/tanzania/dl_catalog_train.csv"),
                          stringsAsFactors = FALSE)
files_should <- c(catalog_train$label, catalog_train$img)
files_have <- file.path("train", list.files(here("results/tanzania/train")))
setdiff(files_should, files_have)

catalog_validate <- read.csv(here("results/tanzania/dl_catalog_valid.csv"),
                             stringsAsFactors = FALSE)
files_should <- c(catalog_validate$label, catalog_validate$img)
files_have <- file.path("validation", list.files(here("results/tanzania/validation")))
setdiff(files_should, files_have)
```

Move images

```{r}
move_train <- lapply(1:nrow(north_train), function(n) {
    things_to <- unlist(north_train[n, c("label", 'img')])
    things_from <- gsub("train", "dl_train", things_to)
    things_to <- file.path(here("results/tanzania"), things_to)
    things_from <- file.path(here("results/north"), things_from)
    file.copy(things_from, things_to)
})

move_valid <- lapply(1:nrow(north_valid), function(n) {
    things_to <- unlist(north_valid[n, c("label", 'img')])
    things_from <- gsub("validation", "dl_valid", things_to)
    things_to <- file.path(here("results/tanzania"), things_to)
    things_from <- file.path(here("results/north"), things_from)
    file.copy(things_from, things_to)
})
```

### Train U-Net model for the whole country

The scripts used for training U-Net are under directory hrlcm.
